---
title: "p8105_hw3_anw2158"
author: "Aung Nay Win"
date: "2023-10-12"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#### Problem 1
## Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

#### Problem 2

```{r}
brfss_df =
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health", 
         response %in% c("Poor", "Fair", "Good", "Very good", "Excellent"))%>%
  mutate(response= factor(response, levels = c("Poor", "Fair", "Good", "Very good"  ,"Excellent"), ordered = TRUE)) %>% 
  arrange(response)
```

States that were observed at 7 or more locations in 2002
```{r}
brfss_df %>% 
  filter(year == 2002) %>%
    group_by(locationabbr) %>%
    summarise(locations = n()) %>%
    filter(locations >= 7)
```

States that were observed at 7 or more locations in 2010
```{r}
brfss_df %>% 
  filter(year == 2010) %>%
    group_by(locationabbr) %>%
    summarise(locations = n()) %>%
    filter(locations >= 7)
```

```{r}
brfss_df %>%
    filter(response == "Excellent") %>%
    group_by(year, locationabbr) %>% 
    summarise(avg_data_value = mean(data_value, na.rm = TRUE)) %>% 
  ggplot(aes(x=year,y=avg_data_value, group=locationabbr, color=locationabbr))+
  geom_line() +
    labs(title = "Average 'Excellent' Responses Over Time by State",
         x = "Year",
         y = "Average Data Value") +
    theme_minimal() +
    theme(legend.position = "none") 
```

```{r}
ny_df= brfss_df %>%
    filter(year %in% c(2006, 2010), 
           locationabbr == "NY", 
           response %in% c("Poor", "Fair", "Good", "Very good", "Excellent"))
ggplot(ny_df, aes(x = data_value, fill = response)) +
    geom_histogram(binwidth = 5, position = "identity", alpha = 0.5) +
    facet_wrap(~ year) +
    labs(title = "Distribution of Data Value in NY State (2006 and 2010)",
         x = "Data Value",
         y = "Count") +
    scale_fill_manual(values = c("Poor" = "red", "Fair" = "orange", "Good" = "yellow", "Very Good" = "green", "Excellent" = "blue")) +
    theme_minimal()
```

#### Problem 3

```{r}
demogrpahic_df = read_csv("data/nhanes_covar.csv", skip = 4) %>% 
  janitor::clean_names()%>% 
  filter(age >= 21) %>% 
  drop_na()

accelarometer_df = read_csv("data/nhanes_accel.csv") %>% 
  janitor::clean_names()

merged_df= demogrpahic_df %>% 
  inner_join(accelarometer_df, by = "seqn") %>% 
  mutate(sex = recode (sex, "1"= "Male", "2"="Female" ), 
    education = recode(education, 
                          "1" = "Less than high school", 
                          "2" = "High school equivalent", 
                          "3" = "More than high school")
    )
```

```{r}
merged_df %>% 
  count(education, sex) %>% 
  pivot_wider(
    names_from = sex,
    values_from = n) %>% 
  arrange(education) %>% 
  knitr::kable(digits = 2)
```

```{r}
ggplot(merged_df, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distributions by Gender and Education",
       x = "Education Level",
       y = "Age") +
  theme_minimal() +
  scale_fill_manual(values = c("Male" = "blue", "Female" = "pink"))
```

From graph, the female is older than male at the education level of high school equivalent. The age of male and female is approximately equal at the education level of less than high school. The male is older than female at the education level of more than high school. 

```{r}
total_activity_df = merged_df %>%
  pivot_longer(cols = starts_with("min"), 
               names_to = "minute", 
               values_to = "activity") %>% 
  group_by(seqn) %>%
  summarise(total_activity = sum(activity, na.rm = TRUE))
```

```{r}
final_df= total_activity_df %>% 
  inner_join(demogrpahic_df, by = "seqn") %>% 
  mutate(sex = recode (sex, "1"= "Male", "2"="Female" ), 
    education = recode(education, 
                          "1" = "Less than high school", 
                          "2" = "High school equivalent", 
                          "3" = "More than high school")
    )

ggplot(final_df, aes(x = age, y = total_activity, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ education, scales = "free_y") +
  labs(title = "Total Activity by Age, Gender, and Education Level",
       x = "Age",
       y = "Total Activity") +
  theme_minimal() +
  scale_color_manual(values = c("Male" = "blue", "Female" = "pink"))
```

In the high school equivalent education plot, both males and females show a decrease in total activity as they age. Females have higher activity levels throughout adulthood compared to males in this education category. For the less than high school education plot, both genders experience a decrease in total activity as they age. There's a notable crossover at age 40 where males, initially less active, become more active than females. For the third plot which is more than high school education, the pattern of decreasing activity with age persists. Again, females exhibit higher activity levels compared to males, similar to the first category. 

```{r}
long_df= merged_df %>%
  pivot_longer(cols = starts_with("min"), 
               names_to = "minute", 
               values_to = "activity") %>%
  mutate(hour = as.numeric(str_replace(minute, "min", "")) / 60)

ggplot(long_df, aes(x = hour, y = activity, color = sex)) +
  geom_point(alpha=.5, size=.5) +
  geom_smooth(se = FALSE) +
  facet_wrap(~ education) +  
  labs(
    title = "24-hour Activity Time Courses by Education Level",
    x = "Hour of the Day",
    y = "Activity Level",
    color = "Sex"
  ) +
  scale_x_continuous(breaks = seq(0, 24, 4)) + 
  theme_minimal()
```

From the three graphs, the activity level of both male and female is the lowest during 4 hr of the day and the activity level reaches at the peak around 8 hr of the day and decreases after that. Moreover, the female activilty level is higher than male one based on the three plots. 